---
title: "Distributions homework"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# MVP

```{r}

library(tidyverse)
library(janitor)

```


*Question 1.*
Load the data, clean_names() and explore it.

*Answer 1.*

```{r}

phones_data <- read_csv("data/20190928-items.csv") %>% 
  clean_names()

glimpse(phones_data)

```
*Question 2.*
Find the brand with the highest numbers of phones in the dataset.

*Answer 2.*

```{r}

phones_data %>% 
  group_by(brand) %>% 
  summarise(n_phones = n()) %>% 
  slice_max(n_phones)

```
Samsung is the brand with the highest number of phones (397) in the dataset.

*Question 3.*
For your top brand, plot the distribution of phone ratings as a probability density, overlaying a fitted normal distribution. Do you think the normal distribution provides a good model of these ratings?

```{r}

samsung_ratings <- phones_data %>%
  filter(brand == "Samsung") %>%
  select(rating)

# getting the mean and sd for Samsung ratings

samsung_summary <- phones_data %>% 
  filter(brand == "Samsung") %>% 
  summarise(
    num = n(),
    mean = mean(rating),
    sd = sd(rating)
  )

# visualising the distribution of Samsung ratings

samsung_ratings %>% 
  ggplot() +
  aes(x = rating) +
  geom_histogram(aes(y = after_stat(density)),
                 col = "white") +
  stat_function(
    fun = dnorm, 
    args = list(
      mean = samsung_summary$mean,
      sd = samsung_summary$sd
    ),
    col = "darkblue"
)

```

The normal distribution maps well to our histogram, suggesting
that it does show a good model of the Samsung ratings. 


*Question 4.*
We can be more quantitative in comparing the distribution of top brand ratings with a normal distribution. You calculated the mean() and sd() of your distribution in an earlier question. Use these values to calculate the proportions of ratings within one-, two- and three standard deviations of the mean. Compare these proportions with those you expect for a normal distribution.

```{r}

one_sd <- samsung_ratings %>%
  filter(rating >= samsung_summary$mean - samsung_summary$sd) %>%
  filter(rating <= samsung_summary$mean + samsung_summary$sd) %>%
  summarise(prop_within_1sd = n() / nrow(samsung_ratings))

one_sd

two_sd <- samsung_ratings %>%
  filter(rating >= samsung_summary$mean - 2*(samsung_summary$sd)) %>%
  filter(rating <= samsung_summary$mean + 2*(samsung_summary$sd)) %>%
  summarise(prop_within_2sd = n() / nrow(samsung_ratings))

two_sd

three_sd <- samsung_ratings %>%
  filter(rating >= samsung_summary$mean - 3*(samsung_summary$sd)) %>%
  filter(rating <= samsung_summary$mean + 3*(samsung_summary$sd)) %>%
  summarise(prop_within_3sd = n() / nrow(samsung_ratings))

three_sd




```

For a normal distribution, expect 68%
 of values to fall within one standard deviation of the mean. Here, 79%
 of ratings are within this range, so our distribution is slightly too central.

Adding another standard deviation to either side should increase the proportion of values by 27%
 to 95%
 overall in a normal distribution. Our distribution adds 13%
, going to 92%
 overall, so there are too few values in the range from one- to two standard deviations.

Finally, in a normal distribution, adding a further standard deviation to either side should increase the proportion of values by 4.7%
 to 99.7%
 overall. Our distribution adds 6%
 to 98%
, so this is pretty close.







