---
title: "Extension"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}

library(tidyverse)
library(arules)
library(arulesViz)
transactions <- read_csv("data/online_retail_subset.csv")


```
For the first section we are interested in the purchase of two particular items:

item A - ‘HEART OF WICKER SMALL’ (StockCode 22469)
item B - ‘LARGE CAKE TOWEL PINK SPOTS’ (StockCode 21110)

**Question 1.**
Calculate the support for item A (this will be the support for a single item)

**Answer 1** 

```{r}

(support_a <- transactions %>% 
  filter(StockCode == 22469) %>% 
  summarise(support = n()/nrow(transactions)))

```
The support for item A is 0.003.

**Question 2:**

Calculate the support and confidence for rule (A→B). 

**Answer 2**

```{r}

(a_and_b_bought_together <- transactions %>% 
   filter(StockCode %in% c(22469, 21110)) %>% 
   summarise(no_of_invoice = duplicated(InvoiceNo)) %>%
   filter(no_of_invoice == TRUE) %>% 
   count())

(support_a_b <- a_and_b_bought_together / nrow(transactions))

(confidence_a_b <- support_a_b / support_a)

          
```

The support for A→B is 0.00023
The confidence for A→B is 0.06

**Question 3**

Calculate the lift for (A→B)

**Question 3**

```{r}

(support_b <- transactions %>% 
  filter(StockCode == 21110) %>% 
  summarise(support = n()/nrow(transactions)))


(lift_a_b <- support_a_b / (support_a * support_b))


```
The lift (A→B) is 136.36. We can therefore say that A and B are often purchased at the same time. However, as the support is so small, these transactions do not happen often.


```{r}

transactions_reformat <- transactions %>%
  select(InvoiceNo, Description) %>%
  na.omit()

write_csv(transactions_reformat, "transactions_reformat.csv")

apriori_format <- read.transactions("transactions_reformat.csv", 
                                    format = "single", 
                                    sep = ",", 
                                    header = TRUE, 
                                    cols = c("InvoiceNo", "Description"))

```






