---
title: "Probability homework"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---

# 1.1. The Titanic

```{r}

library(gtools)
library(tidyverse)
library(janitor)
source("prob.R")

titanic <- read_csv("data/titanic.csv")

```


```{r}

titanic %>% 
  distinct(class)

crew_list <- c("engineering crew", "victualling crew", "restaurant staff", "deck crew")

# adding crew column

titanic <- titanic %>% 
  mutate(crew = if_else(
    class %in% crew_list, "TRUE", 
    "FALSE")
  )

```

```{r}

contingency_table <- titanic %>% 
  tabyl(survived, crew)

contingency_table

prob_table <- contingency_table %>% 
  adorn_totals(where = c("row", "col")) %>% 
  adorn_percentages(denominator = "all")

prob_table

```

Calculate the conditional probabilities prob(survived∣crew)
and prob(survived∣passenger)to settle this dispute.

*Step 1: calculating conditional probability of prob(survived | crew)*

p(survived GIVEN crew) = p(survived AND crew) / p(crew)

```{r}



p_survived_and_crew <- titanic %>% 
  filter(survived == "yes",
         crew == TRUE) %>% 
  summarise(p = n()/nrow(titanic))

p_crew <- titanic %>% 
  filter(crew == TRUE) %>% 
  summarise(p = n()/nrow(titanic))

p_survived_given_crew <- p_survived_and_crew / p_crew

p_survived_given_crew


```

The probability of surviving if a crew member is 0.2370787

*Step 2: calculating conditional probability of prob(survived | passenger)*

p(survived GIVEN passenger) = p(survived AND passenger) / p(passenger)

```{r}

p_survived_and_passenger <- titanic %>% 
  filter(survived == "yes",
         crew == FALSE) %>% 
  summarise(p = n()/nrow(titanic))

p_passenger <- titanic %>% 
  filter(crew == FALSE) %>% 
  summarise(p = n()/nrow(titanic))

p_survived_given_passenger <- p_survived_and_passenger / p_passenger

p_survived_given_passenger

```

The probability of surviving if a passenger member is 0.3796507

Our results show that there was a slightly higher rate of survival for passengers
than for crew members. 

# 1.2 Probability Consolidation Questions

*Question 1*

If I roll a six-sided dice once, what is the probability of getting a number less than 5?
 
*Answer 1*

```{r}

roll_die_outcomes <- rolldie(1) %>% 
  summarise(outcomes = n())


roll_die_less_than_5 <- rolldie(1) %>% 
  filter(X1 < 5) %>% 
  summarise(less_than_5_outcomes = n())

p_less_than_5 <- roll_die_less_than_5 / roll_die_outcomes

p_less_than_5


```

The probability of rolling one six sided die and getting a number less than 5
is 0.66666 (or 2/3)


*Question 2*

How many outcomes are there in the sample space of an experiment in which we 
toss 3 dice?

*Answer 2* 

```{r}

roll_3_dice_outcomes <- rolldie(3) %>% 
  summarise(outcomes = n())


roll_3_dice_outcomes 

```

The are 216 possible outcomes in an experiment where we toss 3 dice.


*Question 3*

In an experiment where we roll four dice:

i. how many outcomes are in the event where the total of the four dice equals 5, 
6 or 7?

ii. what is the probability of this event?

iii. what assumptions do we apply in answering part ii?

*Answer 3.i*

```{r}

roll_4_dice_total_5_6_7 <- rolldie(4) %>% 
  mutate(roll_total = X1+X2+X3+X4) %>% 
  filter(roll_total %in% c(5, 6, 7)) %>% 
  summarise(outcomes = n()) %>% 
  pull()

roll_4_dice_total_5_6_7

```

There 34 possible outcomes that total 5, 6 or 7 when rolling 4 dice. 

*Answer 3.ii*

```{r}

roll_4_dice_outcomes <- rolldie(4) %>% 
  summarise(outcomes = n()) %>% 
  pull()


p_roll_total_5_6_7 <- roll_4_dice_total_5_6_7 / roll_4_dice_outcomes

p_roll_total_5_6_7
```

The probability of rolling 4 dice and the resulting outcomes totalling 5, 6 or 
7 is 0.02623457


*Answer 3.iii*

We assume that each roll of the dice are completely independent of each other.


*Question 4*

Five chairs are placed in a row in a room, along with five people of differing height.
The people sit down randomly on the chairs. What is the probability that they sit 
in order of decreasing height, left to right?

*Answer 4*

```{r}

n_height_combinations <- factorial(5)

# number of height combinations is 120. The permutation 1, 2, 3, 4, 5 can only occur
# once. The probability can be found by:

p_height_order <- 1/n_height_combinations

p_height_order

```

The probability that the 5 people sit in order of decreasing height is 0.008333

*Question 5*

I roll six dice:

i. what is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the dice,
left to right?
ii. what is the probability I see one each of the numbers 6, 5, 4, 3, 2 and 1 on the 
dice regardless of order?

*Answer 5.i*


```{r}

roll_6_dice <- rolldie(6) 

roll_6_dice_outcomes <- rolldie(6) %>% 
  summarise(n()) %>% 
  pull()

# 46656 different outcomes when rolling 6 dice

# we know that the order 6 5 4 3 2 1 is a permutation that can only occur once,
# but let's double check anyway:

order_654321 <- roll_6_dice %>% 
  filter(X1 == 6,
         X2 == 5,
         X3 == 4,
         X4 == 3,
         X5 == 2,
         X6 == 1) %>% 
  summarise(n()) %>% 
  pull()

p_654321 <- order_654321 / roll_6_dice_outcomes

p_654321


```

The probability of the outcome being the numbers 6, 5, 4, 3, 2 and 1 in order on 
the dice, left to right is 2.143347e-05

*Answer 5.ii*


```{r}

# n_of_outcomes of rolling the a combination of 1, 2, 3, 4, 5, 6
# n = 6, r = 6 
# formula = n! / (n-r)!.

no_of_outcomes <- factorial(6) / factorial(6-6)

p_rolling_123456_comb <- no_of_outcomes / roll_6_dice_outcomes

p_rolling_123456_comb

```

The probability of rolling 1, 2, 3, 4, 5, 6 in any order is 0.0154321


