---
title: "Homework - CIs"
output:
editor_options: 
  chunk_output_type: inline
---

**Task 1.**  

Load the data again, `clean_names()`, and re-familiarise yourself with it

```{r}

library(tidyverse)
library(janitor)
library(infer)

ames <- read_csv("data/ames.csv") %>% 
  clean_names()

```

**Task 2.**  
Investigate the distribution of `lot_area`. Is the distribution roughly normal? If not, what problems do you find?

```{r}

lot_area_plot <- ames %>% 
  ggplot() +
  aes(x = lot_area) +
  geom_histogram(colour = "white",
                 fill = "purple",
                 bins = 50) +
    stat_function(
      fun = dnorm,
      args = list(
        mean = mean(ames$lot_area),
        sd = sd(ames$lot_area)
      ), colour = "red2"
      )


lot_area_plot

```

The distribution does not appear to be normal. The spread of the data, and therefore the scales of the x and y axis make it challenging to fully visualise the distribution. Looking at the dataset we can see that the lot_area data ranges from 1300 to 215245.


**Task 3.**  
Compute and visualise a bootstrap sampling distribution for the `mean(lot_area)` of the sold houses.

```{r}

lot_area_sample <- ames %>% 
  specify(response = lot_area) %>%  
  generate(reps = 5000, type = "bootstrap") %>%
  calculate(stat = "mean")

lot_area_sample

lot_area_sample %>% 
  ggplot() +
  aes(x = stat) +
  geom_histogram(fill = "yellow3", colour = "white")

```



**Task 4.**  
Use your bootstrap distribution to calculate a $95\%$ CI for `mean(lot_area)`, and visualise it on the distribution

```{r}

ci_95 <- lot_area_sample %>% 
  get_ci(level = 0.95, type = "percentile")

ci_95

```

Our 95% CI has a lower value of 9837 and an upper value of 10442.

**Task 5.**  
You would like to know the `mean(lot_area)` of the sold houses with higher confidence. Calculate the $99\%$ CI for this variable (you can re-use your bootstrap distribution from above). Is it narrower or broader than the $95\%$ CI? Does that make sense?

```{r}

ci_99 <- lot_area_sample %>% 
  get_ci(level = 0.99, type = "percentile")

ci_99

```

Our 99% CI has a lower value of 9816 and an upper value of 10553.

This is a wider range than for our 95% CI. This makes sense as we can have higher confidence when looking at a broader range of data. 


**Task 6.**  
Calculate the point estimate of the `mean(lot_area)`

```{r}

mean_lot_area <- lot_area_sample %>% 
  summarise(point_est = mean(stat))

mean_lot_area

```

The point estimate is 10150. This makes sense as it falls neatly within our confidence interval. 

# Extension

**Task 1.**  
Calculate a point estimate and $95\%$ CI for the proportion of houses in the data built before 1920.  Does the number of `reps` you use matter? [Investigate `reps` from $200$ up to $50000$, memory of your laptop permitting].

```{r}

before_1920 <- ames %>% 
  filter(lot_area < 1920) %>% 
  specify(response = lot_area) %>%
  generate(reps = 5000, type = "bootstrap") %>%
  calculate(stat = "mean")

before_1920

before_1920_point_est <- ames %>% 
  filter(lot_area < 1920) %>%
  summarise(point_est = mean(lot_area))

before_1920_point_est

before_1920_ci_95 <- before_1920 %>% 
  get_ci(level = 0.95, type = "percentile")

before_1920_ci_95

before_1920_ci_99 <- before_1920 %>% 
  get_ci(level = 0.99, type = "percentile")

before_1920_ci_99

#ran out of time! 


```
The point estimate for houses built before 1920 is 1659.75.

Our 95% CI has an upper limit of 1619.48 and a lower limit of 1699.68.

Our 99% CI has a lower value of 1606 and an upper value of 1711.










